Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-23 writing/embedded_meta_learning_main.tex
Continuing clarity tweaks

\section{Meta-mappings}
We propose the challenge of meta-mappings, that is, tasks which take tasks as an input or output. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. \par
We argue that these task-to-task meta-mappings are a useful way to think about human-like flexibility. For example, a go-playing system should be able to adapt to a different board size. Go on a large board and go on a small board share a great deal of structure. Humans can exploit this to play on a different board without retraining, but deep learning models generally have no way of achieving this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture this human flexibility. We can adapt to task alterations zero-shot, that is, without seeing a single example from the new task. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
Their are different ways this adaptation can be cued, for example by seeing examples of similar task mappings (``you've played checkers on a regular and half-size board, now try to adapt your go strategy to a smaller board,'') or simply by natural-language instructions (``play go on this smaller board'', see appendix \ref{app_why_meta_mapping} for a discussion of why this is better than e.g. conditioning on a task description.). Achieving the flexibility to adapt to meta-map to new tasks will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

\section{Embedded Meta-Learning (EML) architecture}
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for data and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explain behavior), or meta-map tasks to tasks (change behavior). This is a step closer to the human flexibility. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

More formally, we take a \emph{functional} perspective on learning. A datum can be represented by a constant function which outputs it. (For example, each point in the latent space of an autoencoder can be thought of this way.) This allows us to interpret model inputs or outputs as functions. We can then interpret most machine learning tasks as a mapping of functions to functions. The key insight is that, while these mappings could be over functions that represent data\footnote{Where ``data'' is a quite flexible term. The approach is relatively agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.}, they could also operate on task-mappings themselves. This is related to the idea of a homoiconicity, defined above. Under this perspective, learning tasks, learning to map from language to tasks, and learning to flexibly transform between tasks, are all the same type of problem. \par

@@

\paragraph{Meta-mapping:} However, the system is not limited to simply mapping inputs to outputs. The key advantage of our architecture is that it allows transforming anything that is embedded in $Z$ using the same system. Because tasks are embedded in $Z$, this allows for meta-mappings. For example, suppose we have some embedding $z_{go} \in Z$ for the task of playing go. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from examples of winning and trying-to-lose: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{checkers},z_{checkers,lose}\right), ... \right\}\right)$. Then we can generate a new task embedding $\hat{z}_{new} \in Z$ as:  

@@

Then $\hat{z}_{new}$ could be interpreted as the system's guess at a losing strategy for go. We could then evaluate whether the system loses with the $\hat{z}_{new}$ strategy. All meta-mapping results reported here were evaluated in that way -- evaluating the loss on the target task after transforming the task embedding. \par

>>>>

\section{Meta-mappings}
We propose the challenge of meta-mappings, that is, tasks which take tasks as an input or output. These include mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. \par
We argue that these task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to a variation on that task. For example, go on a large board and go on a small board share a great deal of structure. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way of achieving this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. We can adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
Their are different ways this adaptation can be cued, for example by seeing examples of similar task mappings (``you've played checkers on a regular and half-size board, now try to adapt your go strategy to a smaller board,'') or simply by natural-language instructions (``play go on this smaller board'', see appendix \ref{app_why_meta_mapping} for a comparison to conditioning on a task description.). Achieving the flexibility to adapt to meta-map to new tasks will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

\section{Embedded Meta-Learning (EML) architecture}
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for data and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explain behavior), or meta-map tasks to tasks (change behavior). This is a step closer to the human flexibility. \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

More formally, we take a \emph{functional} perspective on learning. A datum can be represented by a constant function which outputs it. (For example, each point in the latent space of an autoencoder can be thought of this way.) This allows us to interpret model inputs or outputs as functions. We can then interpret most machine learning tasks as a mapping of functions to functions. The key insight is that, while these mappings could be over functions that represent data\footnote{Where ``data'' is a quite flexible term. The approach is relatively agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.}, they could also operate on functions that represent tasks. This is related to the idea of a homoiconicity, defined above. Under this perspective, learning tasks, learning to map from language to tasks, and learning to flexibly transform between tasks, are all the same type of problem. \par

@@

\paragraph{Meta-mapping:} However, the system is not limited to simply mapping inputs to outputs. The key advantage of our architecture is that it allows transforming anything that is embedded in $Z$ using the same system. Because tasks are embedded in $Z$, this allows for meta-mappings. For example, suppose we have an embedding $z_{go} \in Z$ for the task of playing go. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from examples of winning and trying-to-lose various games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{checkers},z_{checkers,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{new} \in Z$:  

@@

This $\hat{z}_{new}$ could be interpreted as the system's guess at a losing strategy for go. We could then evaluate whether the system loses with the $\hat{z}_{new}$ strategy. All meta-mapping results reported here were evaluated in that way -- evaluating the loss on the target task after transforming the task embedding. \par
